---
title: "Untitled"
output: html_document
---
## Basic Binary Classification Modeling

```{r}
iris = iris
summary(iris)
```
Description of distributions:
#there are 3 specicies. length of sepal has mean=5.84, IQR = 1.3, range=3.6
#sepal width mean=3.057, IQR = 0.5, range=2.40
#petal length mean=3.758, IQR = 3.5, range=5.90
#petal width mean=1.199, IQR = 1.5, range=2.40
The outcome metric is `Species`. This is what we will be trying to predict. However, we have only done binary classification in class (i.e. two classes). Thus the first order of business is to drop one class. Let's drop the level "virginica" from the data frame.

```{r}
iris = subset(iris, Species != "virginica")#creates subset of all rows except the ones where species name is virginica
```



```{r}
y = rep(0, 100)
for(i in 1:nrow(iris)) {
  y[i] = as.numeric(iris[i, 5])-1
}
```

Fitting a threshold model for Sepal.length

```{r}
X = iris$Sepal.Length
Xy = cbind(X, y)
Xy = as.data.frame(Xy)
colnames(Xy) = c("Length", "species")
n = nrow(Xy)
num_err_by_par = matrix(NA, nrow = n, ncol = 2)
colnames(num_err_by_par) = c("threshold", "num_errors")
y_check = 0
for(i in 1:n) {
  threshold = Xy$Length[i]
  num_errors = sum((Xy$Length > threshold) != y_check)
  num_err_by_par[i, ] = c(threshold, num_errors)
}
best_row = order(num_err_by_par[, "num_errors"])[1]
num_err_by_par[best_row, "threshold"]
```
Threshold parameter ~ 7

```{r}
summary(iris[iris$Species == "setosa", "Sepal.Length"])
summary(iris[iris$Species == "virginica", "Sepal.Length"])
```


My own Perceptron model 

```{r}
max_iterations = 1000
y = as.numeric((iris$Species))-1  #using same y from question : Now create a vector `y` that is length the number of remaining rows in the data frame whose entries are 0 if "setosa" and 1 if "versicolor".
w_vec = rep(0, 5)#initialize weights
feats = iris[ , 1:4]#set feats to the columns of the iris data (without species col) & with no instances of virginica
feats1 = as.matrix(cbind(1, feats))
for(iter in 1:max_iterations) {
  for(i in 1:nrow(feats1)) {
    x_i = feats1[i, ] #set vector x_1 equal to i'th row containing all its' columns
    y_i = y[i] #set y_i equal to the binary value located at postion i in the vector y
    yhat_i = ifelse(sum(x_i %*% w_vec) > 0, 1, 0)
    w_vec = w_vec + (y_i - yhat_i)*x_i
  }
}
w_vec

```

