
Linear Modelling(ordinary least squares mathematically)

Sample Data
```{r}
n = 20
x = runif(n)
beta_0 = 3
beta_1 = -2
y = beta_0 + beta_1 * x + rnorm(n, mean = 0, sd = 0.33)
```

Solving for the least squares line by computing $b_0$ and $b_1$ _without_ using the functions `cor`, `cov`, `var`, `sd` but instead computing it from the $x$ and $y$ quantities manually.
```{r}
ybar = (1/n) * sum(y)
xbar = 1/n * sum(x)
b_1 = (sum(x*y)-n*xbar*ybar)/(sum(x^2)-n*xbar^2)
b_0 = ybar - b_1*xbar
g = b_0 + b_1*x
```

A test to check if equations above are correct:
```{r}
lm_mod = lm(y ~x)
b_vec = coef(lm_mod)
expect_equal(b_0, as.numeric(b_vec[1]), tol = 1e-4) 
expect_equal(b_1, as.numeric(b_vec[2]), tol = 1e-4)
```



a recreation of Galtons famous experiment
```{r}
pacman::p_load(HistData)
data("Galton")
```
A short description of the data set:
This data set is made up of 928 observations of parent height and their childs height. Parent height is the average of the mother & father height. Female childs height was multipled by 1.08 before it was recorded to account for a difference in sex. Glaton used classes to group the data into, where each class interal has a width of 1.0 inch.

Running a linear model to explain the childrens' height using the parents' height:
```{r}
gal = lm(Galton$child ~ Galton$parent)
coefficients(gal)#returns b0 and b1
summary(gal)$r.squared
summary(gal)$sigma#RMSE
```
Explanation of results:
b_0=23.94
b1=.65 for every addition inch in parents height, we can expect the childrens height to increase by .65 inches
R^2=.21 -the independent variables don't explain the dependent variables well
RMSE=2.24-our model is not very accurate when predicting childrens height

Plot data and best fit line:
```{r}
pacman::p_load(ggplot2)
ggplot(data = Galton, aes(x = parent, child)) + geom_point() + geom_abline(intercept = 23.94, slope = .65)
```

Let's plot (a) the data in the set as black dots, (b) least squares line defined by $b_0$ and $b_1$ in blue, (c) the theoretical line $\beta_0$ and $\beta_1$ if the parent-child height equality held in red and (d) the mean height in green.
```{r}
ggplot(Galton, aes(x = parent, y = child)) + 
  geom_point() + 
  geom_jitter() +
  geom_abline(intercept = 23.94, slope = .65, color = "blue", size = 1) +
  geom_abline(intercept = 0, slope = 1, color = "red", size = 1) +
  geom_abline(intercept = avg_height, slope = 0, color = "darkgreen", size = 1) +
  xlim(63.5, 72.5) + 
  ylim(63.5, 72.5) +
  coord_equal(ratio = 1)
```
This graph shows the conclusion that children of short parents, will be, on average, short and children of tall parents will be, on average, tall.
